---
title: "WUP_04"
author: "Ever De Leon"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS <- read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

The numerical variable is the average score in the WORDSUM test and the categorical variable is the level 
<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We want to predict the WORDSUM scores based on DEGREE.

Data from GSS:

WORDSUM ~ DEGREE

WORDSUM is a numerical variable

DEGREE is a categorical variable

So we use the Numerical~Cat helper file

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=GSS)
```

Based on the boxplot, there are differences in the median values between every degree with LT HIGH SCHOOL being the lowest with a median value of 8 compared to the average score in the WORDSUM test in the GRADUATE with a median of 11.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSS)
```

Based on the graph above: 
BACHELOR has a WORDSUM score mean of 9.204545.  
GRADUATE has a WORDSUM score mean of 10.095238.  
HIGH SCHOOL has a WORDSUM score mean of 8.282353.  
JUNIOR COLLEGE has a WORDSUM score mean of 9.916667.  
LT HIGH SCHOOL has a WORDSUM score mean of 7.750000.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSS)
summary.aov(model1)
```

Since the p-value is 0.000236, which is less than 0.05, there is evidence this is statistically significant.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Our question is whether thre is any difference in between average score in the WORDSUM test and the level of scoring based on DEGREE. According to the boxplot, the median values differed with the lowest being LT HIGH SCHOOL and the highest being GRADUATE. Based on the numerical description graph, it further proves that WORDSUM score medians increase as the DEGREE level increases. Furthermore, the p-value of 0.000236 is less than 0.05. Therefore we can reject the $H_0$ and demonstate that there is a correlation in WORDSUM scores and DEGREE.

